---
title: "Programming assignment 1"
author: "Triin Kask"
date: "31 December 2017"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Programming assignment in Reproducible Research

### First read the file in
```{r read file, echo=TRUE}
unzip("repdata_activity.zip")
data <- read.csv("activity.csv")
```

### Then create a histogram of sum of steps per day
```{r histo file, echo=TRUE, fig.height=6, fig.widht=6}
#Calculate the values
result <- aggregate(data$steps, by=list(data$date), sum)
result$x <- as.numeric(as.character(unlist(result$x)))
result <- na.omit(result)
colnames(result) <- c("Date","Sum_steps")
library(ggplot2)
ggplot(result, aes(x = Date,y = Sum_steps)) + geom_bar(stat = "identity", colour = result$Date, fill = "white" ) + xlab("Date") + ylab("Steps sum count") + ggtitle("Sum of steps per day") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Then show the mean and median of steps per day
```{r histo mean and median, echo=TRUE}
#Calculate the mean and median
mean(result$Sum_steps)
median(result$Sum_steps)
```

### Next calculate the daily activity pattern
```{r activity pattern, echo=TRUE, fig.height=6, fig.widht=6}
#Calculate activity
average <- aggregate(data$steps, by = list(data$interval), mean, na.rm = TRUE)
colnames(average) <- c("step_interval", "steps")

#Plot averages
ggplot(data = average, aes(x = step_interval, y = steps)) + geom_line() + xlab("5-minute interval") + ylab("Average number of steps") + ggtitle("Daily activity pattern")
```

### Find the maximal value
```{r activity max, echo=TRUE}
maximum <- average[which.max(average$steps),]
maximum
```

### Find the total number of NAs in this data.frame
```{r missing values, echo=TRUE}
count <- is.na(data$steps)
table(count)
```

### Fill all the missing values in this dataframe and create a new one
```{r missing values fill, echo=TRUE, fig.height=6, fig.widht=6}
#Fill missing values
data <- read.csv("activity.csv")
for (i in 1:nrow(data)){
    if (is.na(data$steps[i])){
        data$steps[i] <- average$steps[which(data$interval[i] == average$step_interval)]}
}
total_steps <- aggregate(data$steps, by = list(data$date), sum, na.rm = TRUE)
colnames(total_steps) <- c("interval","steps")

#Plot a histogram
ggplot(total_steps, aes(x = interval, y = steps)) + geom_histogram(stat = "identity") + xlab("5-minute interval") + ylab("Number of steps") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("Total steps per day with filled data")
```

### Calculated the mean and median or result
```{r missing mean and result, echo=TRUE}
values <- as.numeric(as.character(unlist(total_steps$steps)))
mean(values)
median(values)
```

### Show differences between weekdays and weekends
```{r weekday and weekend, echo=TRUE, fig.height=6, fig.widht=6}
#New factor - "weekday" and "weekend" as a new column taken from the date
data$weekdays <- weekdays(as.Date(data$date))
data$weekdays <- ifelse(data$weekdays %in% c("Saturday", "Sunday"),"weekend", "weekday")

#Plot a histogram
plotted_data <- data
average_values <- aggregate(steps ~ interval + weekdays, data = plotted_data, mean)
ggplot(average_values, aes(interval, steps)) + geom_line(stat= "identity", color = "blue") + facet_grid(weekdays ~ .) + xlab("5-minute interval") + ylab("Number of steps") + geom_smooth(position = "identity", method = "auto")
```
